---
title: "DMIS and Xiaoping top predictions - similarity analysis"
author: "Robert Allaway, Sage Bionetworks"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

The top compounds selected by each team had binding in kinase assays, but did not have activity in flies. To test some other compounds, we requested the top 100 compounds from each team. One team (Zhaoping) sent us more than 100 because their scoring metric was integer based - eg. after the first few, many compounds had the same ranking - so we received the top 253 and 443 compounds for P1 and P2 from this team, respectively. DMIS-MTD sent us the top 100 for each problem. Note that their method has stochastic steps, so their results may differ from the challenge submission. Unlike the challenge, there were no identical predictions across the two sets. 

 
```{r message=FALSE, warning=FALSE, include=FALSE}
Sys.Date()
library(pheatmap)
library(tidyverse)
library(synapser)
synLogin()

library(rcdk)
library(webchem)

zp_p1 <- read_csv(synGet("syn20803149")$path, skip = 1L) %>% top_n(100,score)
zp_p2 <- read_csv(synGet("syn20803150")$path, skip = 1L) %>% top_n(100,score)

dmis_p1 <- read_tsv(synGet("syn20803147")$path)
dmis_p2 <- read_tsv(synGet("syn20803148")$path)

```

## Problem 1

We processed all of the P1 compounds and clustered the similarity matrix. Here, the rows are DMIS-MTD predictions and the columns are Zhaoping's predictions. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

zp_p1_mols <- parse.smiles(zp_p1$smiles)
zp_p1_fps <- lapply(zp_p1_mols, get.fingerprint, type='standard')

dmis_p1_mols <- sapply(dmis_p1$InChI, cs_inchi_smiles) %>% parse.smiles()
dmis_p1_fps <- lapply(dmis_p1_mols, get.fingerprint, type='standard')

fp.sim_p1 <- fingerprint::fp.sim.matrix(dmis_p1_fps, zp_p1_fps, method='tanimoto')

rownames(fp.sim_p1) <- dmis_p1$`ZINC ID`
colnames(fp.sim_p1) <- zp_p1$zinc_id


pheatmap(fp.sim_p1, 
         color = viridis::plasma(1000, direction = 1, begin = 0, end = 1), 
         show_rownames = F,
         show_colnames = F)

```

Then, we took all interactions where the tanimoto score to any compound in the other dataset was greater than 0.4 (selected simply for ease of visualization) and converted this data to network. Larger weights indicate a larger tanimoto score. Yellow nodes are DMIS, blue nodes are Zhaoping. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(visNetwork)

tidy_sim_p1 <- fp.sim_p1 %>% 
  as_tibble(rownames = NA) %>% 
  rownames_to_column("zp_cmpd") %>% 
  gather(key = "dmis_cmpd", value = "tanimoto", -zp_cmpd) %>% 
  filter(tanimoto > 0.4)

p1_nodes <- select(tidy_sim_p1, -tanimoto) %>% 
  gather(key = "group", value = "id") %>%  
  mutate(label = paste0(group, "_", id)) %>% 
  distinct()

p1_edges <- tidy_sim_p1 %>% 
  set_names(c("from", "to", "value")) 


visNetwork(p1_nodes, p1_edges)

```


## Problem 2

We processed all of the P2 compounds and clustered the similarity matrix. Here, the rows are DMIS-MTD predictions and the columns are Zhaoping's predictions. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
zp_p2_mols <- parse.smiles(zp_p2$smiles)
zp_p2_fps <- lapply(zp_p2_mols, get.fingerprint, type='standard')

dmis_p2_mols <- sapply(dmis_p2$InChI, cs_inchi_smiles) %>% parse.smiles()
dmis_p2_fps <- lapply(dmis_p2_mols, get.fingerprint, type='standard')

fp.sim_p2 <- fingerprint::fp.sim.matrix(dmis_p2_fps, zp_p2_fps, method='tanimoto')

rownames(fp.sim_p2) <- dmis_p2$`ZINC ID`
colnames(fp.sim_p2) <- zp_p2$zinc_id


pheatmap(fp.sim_p2, 
         color = viridis::plasma(1000, direction = 1, begin = 0, end = 1), 
         show_rownames = F,
         show_colnames = F)
```


Then, we took all interactions where the tanimoto score to any compound in the other dataset was greater than 0.4 (selected simply for ease of visualization) and converted this data to network. Larger weights indicate a larger tanimoto score. Yellow nodes are DMIS, blue nodes are Zhaoping. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

tidy_sim_p2 <- fp.sim_p2 %>% 
  as_tibble(rownames = NA) %>% 
  rownames_to_column("zp_cmpd") %>% 
  gather(key = "dmis_cmpd", value = "tanimoto", -zp_cmpd) %>% 
  filter(tanimoto > 0.4)

p2_nodes <- select(tidy_sim_p2, -tanimoto) %>% 
  gather(key = "group", value = "id") %>%  
  mutate(label = paste0(group, "_", id)) %>% 
  distinct()

p2_edges <- tidy_sim_p2 %>% 
  set_names(c("from", "to", "value")) 


visNetwork(p2_nodes, p2_edges)

```
