---
title: "compound scores overview"
author: "Xindi Guo, Kristen Dang"
date: "4/4/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(synapser)
synLogin()
library(gplots)
library(RColorBrewer)
```

## Review of Tanimoto scores

There are two problems in this challenge; each has a set of targets and antitargets. Participants can suggest multiple compounds as solutions for each problem.

For each compound that is submitted, we calculate a Tanimoto score to assess its structural similarity to known small-molecule ligands of the targets.

## Parameters of calculation

- ChEMBL as source of ligands
- Have Kd or IC50 < 10 uM, **per each** target
- Tanimoto calculation on Morgan Fingerprint, radius 3 (ECFP6-like) is used


## Compounds passing Tanimoto

```{r include=FALSE}
allCompounds = as.data.frame(synTableQuery('select * from syn12137044',includeRowIdAndRowVersion = FALSE))
problemInfo = as.data.frame(synTableQuery('select * from syn12081609',includeRowIdAndRowVersion = FALSE))


problemInfo$target[which(problemInfo$target == "Aurora Kinase A")] = "AURKA"

targets = union(grep(colnames(allCompounds),pattern = "CHEMB"), grep (colnames(allCompounds), pattern="pass"))
for (i in targets){ 
  toReplace = which(allCompounds[,i] == 'true')
  x = replace(allCompounds[,i],toReplace,TRUE)
  allCompounds[,i] = x
  }
for (i in targets){ 
  toReplace = which(allCompounds[,i] == 'false')
  x = replace(allCompounds[,i],toReplace,FALSE)
  allCompounds[,i] = as.logical(x)
}

prob1ReqBind = which(problemInfo$bind == 'true' & problemInfo$required == 'true' & problemInfo$problem == 1)
prob1AllBind = which(problemInfo$bind == 'true' & problemInfo$problem == 1)
prob2ReqBind = which(problemInfo$bind == 'true' & problemInfo$required == 'true' & problemInfo$problem == 2)
prob2AllBind = which(problemInfo$bind == 'true' & problemInfo$problem == 2)


# Does it pass for BINDING required for problem 1 (optional and anti-targets not considered)
colToCheck = which(colnames(allCompounds) %in% c(problemInfo$chembl_id[prob1ReqBind]))
rowProb1 = which(allCompounds$problem == 1)
passProbReq = allCompounds[which(allCompounds$problem == 1),colToCheck] == TRUE

passBindReq = rep(NA,nrow(allCompounds))
passBindReq[rowProb1] = passProbReq


# Does it pass for BINDING required for problem 2 (optional and anti-targets not considered)
colToCheck = which(colnames(allCompounds) %in% c(problemInfo$chembl_id[prob2ReqBind]))
rowProb2 = which(allCompounds$problem == 2)
passProbReq2 = allCompounds[which(allCompounds$problem == 2),colToCheck] == TRUE

# Check that will not overwrite any problem 1 values
intersect(rowProb1,rowProb2)
passBindReq[rowProb2] = passProbReq2[,1] & passProbReq2[,2]



# Does it pass for BINDING required and optional for problem 1 (anti-targets not considered)
colToCheck = which(colnames(allCompounds) %in% c(problemInfo$chembl_id[prob1AllBind]))
rowProb1 = which(allCompounds$problem == 1)
passProbAll = allCompounds[which(allCompounds$problem == 1),colToCheck] == TRUE

passBindAll = rep(NA,nrow(allCompounds))
passBindAll[rowProb1] = passProbAll[,1] & passProbAll[,2] & passProbAll[,3] & passProbAll[,4]


# Does it pass for BINDING Alluired and optional for problem 2 (anti-targets not considered)
colToCheck = which(colnames(allCompounds) %in% c(problemInfo$chembl_id[prob2AllBind]))
rowProb2 = which(allCompounds$problem == 2)
passProbAll2 = allCompounds[which(allCompounds$problem == 2),colToCheck] == TRUE

# Check that will not overwrite any problem 1 values
intersect(rowProb1,rowProb2)
passBindAll[rowProb2] = passProbAll2[,1] & passProbAll2[,2] & passProbAll2[,3] & passProbAll2[,4]


allCompounds$passBindReq = passBindReq
allCompounds$passBindAll = passBindAll



## How many compounds pass Tanimoto?

counts = matrix(NA,nrow = 4, ncol = 2)
colnames(counts) = c("pass", "fail")
rownames(counts) = c("required targets 1", "all targets 1", "required targets 2", "all targets 2")
sumCols = grep(colnames(allCompounds),pattern = "passBindReq")
counts[1,1] = sum(allCompounds[rowProb1,sumCols])
counts[3,1] = sum(allCompounds[rowProb2,sumCols])

sumCols = grep(colnames(allCompounds),pattern = "passBindAll")
counts[2,1] = sum(allCompounds[rowProb1,sumCols])
counts[4,1] = sum(allCompounds[rowProb2,sumCols])

# Count fail
counts[1,2] = length(rowProb1) - counts[1,1]
counts[3,2] = length(rowProb2) - counts[3,1]

counts[2,2] = length(rowProb1) - counts[2,1]
counts[4,2] = length(rowProb2) - counts[4,1]
```

```{r Tanimoto, echo = FALSE}
op = par()
par(mar = c(10,4,4,2))
barplot(t(counts),beside = FALSE,legend.text = colnames(counts), las = 2, ylab = "number of submitted compounds", main = "Tanimoto score results")
par(op)
```

## Average pass rate per target?

```{r pertarget, echo=FALSE}
targets = grep(colnames(allCompounds),pattern = "CHEMB")
row1Pass = colSums(allCompounds[rowProb1,targets])
row2Pass = colSums(allCompounds[rowProb2,targets])

boxplot(row1Pass/length(rowProb1), row2Pass/length(rowProb2), ylab = "Fraction of compounds passing Tanimoto", main = "Solution uniqueness per target", names = c("problem 1", "problem 2"))
```

## Median Tanimoto score per target
```{r medScore, echo=FALSE}
medTanimoto = as.data.frame(synTableQuery('select * from syn12137045',includeRowIdAndRowVersion = FALSE))
# Make problem 1 matrix
targetProb1 = which(colnames(medTanimoto) %in% c(problemInfo$chembl_id[prob1AllBind]))
prob1Tan = medTanimoto[rowProb1,targetProb1]
colnames(prob1Tan) = problemInfo$target[match(colnames(prob1Tan), problemInfo$chembl_id)]
# Make problem 2 matrix
targetProb2 = which(colnames(medTanimoto) %in% c(problemInfo$chembl_id[prob2AllBind]))
prob2Tan = medTanimoto[rowProb2,targetProb2]
colnames(prob2Tan) = problemInfo$target[match(colnames(prob2Tan), problemInfo$chembl_id)]

targetsT = grep(colnames(medTanimoto),pattern = "CHEMB")

# Tanimoto per target
op = par()
par(mfrow = c(1,2))
boxplot(prob1Tan, las = 2, ylab = "median Tanimoto score per compound", main = "problem 1", ylim = range(min(medTanimoto[,targetsT]), max(medTanimoto[,targetsT])))
boxplot(prob2Tan, las = 2, ylab = "median Tanimoto score per compound", main = "problem 2", ylim = range(min(medTanimoto[,targetsT]), max(medTanimoto[,targetsT])))
par(op)
```


## Median Tanimoto score per team
```{r medTeam, echo=FALSE}

# Tanimoto per team
op = par()
par(mfrow = c(1,2))

prob1wide = reshape(medTanimoto[rowProb1,c(1,2,3,targetProb1)], v.names = colnames(medTanimoto)[targetProb1], idvar = "submission_id", direction = "wide", timevar = "solution", drop = "problem")
prob2wide = reshape(medTanimoto[rowProb2,c(1,2,3,targetProb2)], v.names = colnames(medTanimoto)[targetProb2], idvar = "submission_id", direction = "wide", timevar = "solution", drop = "problem")

boxplot(t(prob1wide[,2:ncol(prob1wide)]), names = prob1wide[,1], las = 2, ylab = "median Tanimoto score per compound", main = "problem 1")
boxplot(t(prob2wide[,2:ncol(prob2wide)]), names = prob2wide[,1], las = 2, ylab = "median Tanimoto score per compound", main = "problem 2")
par(op)

```

## Compound and target clustering  
```{r heat1, echo=FALSE}
plotcols = brewer.pal(n = length(unique(medTanimoto$submission_id)),name = "Set3")

y = data.matrix(medTanimoto[rowProb1,targetProb1])
colnames(y) = problemInfo$target[match(colnames(y), problemInfo$chembl_id)]

heatmap.2(t(y),scale = "row",na.rm = TRUE, ColSideColors = plotcols[as.numeric(as.factor(medTanimoto$submission_id[rowProb1]))],labCol = "",trace = "none", main = "problem 1",cexRow = 1.1,col = brewer.pal(11,"BrBG"))

```

## Compound and target clustering 
```{r heat2, echo=FALSE}
y = data.matrix(medTanimoto[rowProb2,targetProb2])
colnames(y) = problemInfo$target[match(colnames(y), problemInfo$chembl_id)]
heatmap.2(t(y),scale = "row",na.rm = TRUE, ColSideColors = plotcols[as.numeric(as.factor(medTanimoto$submission_id[rowProb2]))], labCol = "",trace = "none", main = "problem 2",cexRow = 1.1,,col = brewer.pal(11,"BrBG"))
```


## How many compounds pass Lipinski and polar surface area?

```{r lipinski, echo=FALSE}
lipPSA = as.data.frame(synTableQuery('select submission_id, problem, solution, pass_lipinski, pass_polar_surface_area from syn12063027', includeRowIdAndRowVersion = FALSE))

targets = grep (colnames(lipPSA), pattern="pass")
for (i in targets){ 
  toReplace = which(lipPSA[,i] == 'true')
  x = replace(lipPSA[,i],toReplace,TRUE)
  lipPSA[,i] = x
}
for (i in targets){ 
  toReplace = which(lipPSA[,i] == 'false')
  x = replace(lipPSA[,i],toReplace,FALSE)
  lipPSA[,i] = as.logical(x)
}

counts = matrix(NA,nrow = 2, ncol = 2)
colnames(counts) = c("pass", "fail")
rownames(counts) = c("problem 1", "problem 2")
rowProb1 = which(lipPSA$problem == 1)
rowProb2 = which(lipPSA$problem == 2)
sumCols = grep(colnames(lipPSA),pattern = "lipinski")
counts[1,1] = sum(lipPSA[rowProb1,sumCols])
counts[2,1] = sum(lipPSA[rowProb2,sumCols])

# Count fail
counts[1,2] = length(rowProb1) - counts[1,1]
counts[2,2] = length(rowProb2) - counts[2,1]

op = par()
par(mar = c(10,4,4,2))
par(mfrow = c(1,2))
barplot(t(counts),beside = FALSE,legend.text = colnames(counts), las = 2, ylab = "number of submitted compounds", main = "Lipinski rule results")


## How many compounds pass polar surface area criterion?

counts = matrix(NA,nrow = 1, ncol = 2)
colnames(counts) = c("pass", "fail")
rownames(counts) = c("problem 2")
rowProb2 = which(lipPSA$problem == 2)
sumCols = grep(colnames(lipPSA),pattern = "polar")
counts[1,1] = sum(lipPSA[rowProb1,sumCols])

# Count fail
counts[1,2] = length(rowProb1) - counts[1,1]

barplot(t(counts),beside = FALSE,legend.text = colnames(counts), las = 2, ylab = "number of submitted compounds", main = "PSA rule results")
par(op)
```

## Compounds passing all criteria

```{r final, echo = FALSE}
targets = union(union(grep(colnames(allCompounds),pattern = "CHEMB"), grep(colnames(allCompounds), pattern="for")), grep(colnames(allCompounds), pattern = "smiles"))
combinedData = merge(lipPSA, allCompounds[,-targets])

par(mfrow = c(1,2))
barplot(table(rowSums(combinedData[rowProb1,c(4,6)])), col = rainbow(5), main = "Success rate - problem 1", ylab = "Number of compounds", xlab = "Number of criteria met")
barplot(table(rowSums(combinedData[rowProb2,c(4,5,6)])), col = rainbow(5), main = "Success rate  - problem 2", ylab = "Number of compounds", xlab = "Number of criteria met")
```